Lyman-alpha spectra from rotating outflows

### 2018-03-26 paper/paper.tex
making paper plots in pdf format

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
distributions of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these two
features impact the morphology of the Lyman-$\alpha$ emission line.
One of the most notorious results is that rotation increases the
detected flux at the line's center in comparison with the
non-rotational case. 
Rotation also induces a line broadening and a dependency with the
viewing angle.
We also present a semi-analytic model that modifies the spectra of
non-rotational that manages to reproduce the main features of these
Monte Carlo simulations. 
As an application example we contraint the rotation parameters to the
observed Lyman-$\alpha$ profile of a high redshift LAE, providing the
first a constraint on the rotational velocity of a high redshift
galaxy based on its integrated \lya emission.
\end{abstract}

@@

These kinematic maps show that LAEs have two  overimposed velocity
patterns: coherent rotation plus and a random motions.
Depending on the system under study, one of the two features can be
more important, or they can be found in equal measure.  
These kinematic features play an important role in shaping the \lya. 
Models of the \lya have also shown that there is clear signal for
outflowing gas most probably produced by star formation activity.
The characteristic signal for these outflows is the asymmetrical \lya
linea with a more pronounced read peak.
The effects of either outflows or rotation on the \lya line have been
already explored.

Here we present for the first time a study of the joint effects of
galaxy outflows and rotation, two of the three major kinematic
components expected in LAEs. 
We study a simplified geometrical configuration corresponding to a
spherical gas cloud with symmetrical radial outflows and a rotation
profile corresponding to a solid body.
We base our modeling on a Monte-Carlo radiative transfer code called
CLARA (Code for Lyman Alpha Radiation Analysis) presented for the
first time by \cite{CLARA}.

Besides modeling the impact of joint rotation and outflows, we also
check to what extent the analytical model presented by
\cite{Garavito14} to explain the effects of rotation can also be
applied in our case.
We show how solid body rotation effects can be modelled by
postprocessing the results of outflow only simulation.

in the results of other radiative transfer model in the case of large
optical depths. 
In this case it is a good approximation to Doppler
boost the results of the model without rotation.

@@

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_6}
  \end{center}
  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=90^\circ$. 
  	We vary \vrot increasing from left to right and \vout increasing from top to 
  	bottom. The legend represents the \lya line obtained if there is no rotation 
  	(in thin line), if there is a radiative transfer of rotation and outflows 
  	(thick and clear line), and if there is a radiative transfer of only outflows,
  	but also a Doppler shift from the rotational velocity (thick and dark line). 
	    \label{fig:doppler_shift}}
\end{figure*}

@@

% Esta figura deberia tener 2 x 2 vinetas. vout=50.
% cada panel tiene una curva para theta=0, 45, 90.
% | tau=1E5, vrot=50, | tau=1E5, vrot=100 |
% | tau=1E6, vrot=50, | tau=1E6, vrot=100 | 
%\begin{figure*}
%  \begin{center}
%    \includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation_small}
%  \end{center}
%  \caption{\textbf{Spectra for fixed outflows velocity:} We fixed $\vout=50$\kms. 
%    We vary \vrot increasing from left to right, \tauh increasing from top to 
%    bottom, and $\theta$ according to the legend.
%    \label{fig:varying_rotation_small}}
%\end{figure*}

@@

% Esta figura deberia tener 2 x 2 vinetas. vrot=50.
% cada panel tiene una curva para theta=0, 45, 90.
% | tau=1E5, vout=25, | tau=1E5, vout=75 |
% | tau=1E6, vout=25, | tau=1E6, vout=75 |
%\begin{figure*}
%\begin{center}
%\includegraphics[width=0.90\textwidth]{./figures/results/varying_outflow_small}
%\end{center}
%\caption{\textbf{Spectra for fixed rotational velocity:} We fixed $\vrot=50$\kms.  We vary \vout in%creasing from left to right, \tauh increasing from top to 
			 bottom, and $\theta$ according to the legend.
%\label{fig:varying_outflow_small}}
%\end{figure*}

@@

% Esta figura deberia tener 3 x 2 viñetas. tau=10E6.
% cada panel tiene una curva para no rotation, rotation (radiative transfer)
% y rotation (doppler)
% |vrot=50, vout=5,  | vrot=100, vout=5,  |
% |vrot=50, vout=25, | vrot=100, vout=25, |
% |vrot=50, vout=50, | vrot=100, vout=50, |

@@

\begin{figure*}
\begin{center}
\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout5_logtau7}
\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout25_logtau7}
\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout50_logtau7}
\end{center}
\caption{\textbf{Standard Deviation for RT and DS:} The triangles represent the Radiative 
  Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
  \label{fig:standard_deviation}}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
  \end{center}
  \caption{\textbf{Skewness for RT and DS:} The triangles represent the Radiative 
    Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
    \label{fig:skewness}}
\end{figure*}

@@

% Esta figura deberia tener en el eje x theta y en el eje y intensity of valley
% cada panel tiene una curva para logtau = 5, 6, 7.
\begin{figure*}
\begin{center}
\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
\end{center}
\caption{\textbf{Valley Intensity:} We show for each \tauh the dependency that
		the viewing angle $\theta$ has on the line's the valley intensity. The larger 
		$\cos{\theta}$ the lower is the valley intensity, so the higher $\theta$, the 
		higher the valley intensity. We fixed $\vout=5$\kms and $\vrot=100$\kms. 
		\label{fig:valley_intensity}}
\end{figure*}

>>>>

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
distributions of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these two
features impact the morphology of the Lyman-$\alpha$ emission line.
One of the most notorious results is that rotation increases the
detected flux at the line's center in comparison with the
non-rotational case. 
Rotation also induces a line broadening and a dependency with the
viewing angle.
We present a semi-analytic model that modifies the spectra of
non-rotational that manages to reproduce the main features of these
Monte Carlo simulations. 
As an application example we contraint the rotation parameters to the
observed Lyman-$\alpha$ profile of a high redshift LAE, providing the
first a constraint on the rotational velocity of a high redshift
galaxy based on its integrated \lya emission.
\end{abstract}

@@

These kinematic maps show that LAEs have two  overimposed velocity
patterns: coherent rotation plus and a random motions around the mean
value provided by rotation.  
Depending on the system under study, one of the two features can be
more important, or they can be found in equal measure.  
These kinematic features play an important role in shaping the \lya.

Models of the \lya have also shown that there is clear signal for
outflowing gas most probably produced by star formation activity.
The characteristic signal for these outflows is the asymmetrical \lya
linea with a more pronounced read peak.

The effects of either outflows or rotation on the \lya line have been
already explored.
Here we present for the first time a study of the joint effects of
galaxy outflows and rotation, two of the three major kinematic
components expected in LAEs.

We study a simplified geometrical configuration corresponding to a
spherical gas cloud with symmetrical radial outflows and a rotation
profile corresponding to a solid body.
We base our modeling on a Monte-Carlo radiative transfer code called
CLARA (Code for Lyman Alpha Radiation Analysis) presented for the
first time by \cite{CLARA}.

Besides modeling the impact of joint rotation and outflows, we also
check to what extent the analytical model presented by
\cite{Garavito14} to explain the effects of rotation can also be
applied in our case.
We show that solid body rotation effects can be included
in the results of other radiative transfer model in the case of large
optical depths. In this case it is a good approximation to Doppler
boost the results of the model without rotation.

@@

@@

% Esta figura deberia tener 2 x 2 vinetas. vout=50.
% cada panel tiene una curva para theta=0, 45, 90.
% | tau=1E5, vrot=50, | tau=1E5, vrot=100 |
% | tau=1E6, vrot=50, | tau=1E6, vrot=100 | 
\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation_small}
	\end{center}
	\caption{\textbf{Spectra for fixed outflows velocity:} We fixed $\vout=50$\kms. 
			 We vary \vrot increasing from left to right, \tauh increasing from top to 
			 bottom, and $\theta$ according to the legend.
	\label{fig:varying_rotation_small}}
\end{figure*}

@@

% Esta figura deberia tener 2 x 2 vinetas. vrot=50.
% cada panel tiene una curva para theta=0, 45, 90.
% | tau=1E5, vout=25, | tau=1E5, vout=75 |
% | tau=1E6, vout=25, | tau=1E6, vout=75 |
\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/results/varying_outflow_small}
	\end{center}
	\caption{\textbf{Spectra for fixed rotational velocity:} We fixed $\vrot=50$\kms. 
			 We vary \vout increasing from left to right, \tauh increasing from top to 
			 bottom, and $\theta$ according to the legend.
		\label{fig:varying_outflow_small}}
\end{figure*}

@@

% Esta figura deberia tener 3 x 2 viñetas. tau=10E6.
% cada panel tiene una curva para no rotation, rotation (radiative transfer)
% y rotation (doppler)
% |vrot=50, vout=5,  | vrot=100, vout=5,  |
% |vrot=50, vout=25, | vrot=100, vout=25, |
% |vrot=50, vout=50, | vrot=100, vout=50, |
\begin{figure*}
  \begin{center}
    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_6}
  \end{center}
  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=90^\circ$. 
  	We vary \vrot increasing from left to right and \vout increasing from top to 
  	bottom. The legend represents the \lya line obtained if there is no rotation 
  	(in thin line), if there is a radiative transfer of rotation and outflows 
  	(thick and clear line), and if there is a radiative transfer of only outflows,
  	but also a Doppler shift from the rotational velocity (thick and dark line). 
	    \label{fig:doppler_shift}}
\end{figure*}

@@

\begin{figure}
	\begin{center}
		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout5_logtau7}
		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout25_logtau7}
		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout50_logtau7}
	\end{center}
	\caption{\textbf{Mean line for RT and DS:} The triangles represent the Radiative 
		Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
		\label{fig:mean_line}}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout5_logtau7}
		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout25_logtau7}
		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout50_logtau7}
	\end{center}
	\caption{\textbf{Standard Deviation for RT and DS:} The triangles represent the Radiative 
		Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
		\label{fig:standard_deviation}}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout5_logtau7}
		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout25_logtau7}
		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout50_logtau7}
	\end{center}
	\caption{\textbf{Skewness for RT and DS:} The triangles represent the Radiative 
		Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
		\label{fig:skewness}}
\end{figure}

@@

% Esta figura deberia tener en el eje x theta y en el eje y intensity of valley
% cada panel tiene una curva para logtau = 5, 6, 7.
\begin{figure}
	\begin{center}
		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
	\end{center}
	\caption{\textbf{Valley Intensity:} We show for each \tauh the dependency that
		the viewing angle $\theta$ has on the line's the valley intensity. The larger 
		$\cos{\theta}$ the lower is the valley intensity, so the higher $\theta$, the 
		higher the valley intensity. We fixed $\vout=5$\kms and $\vrot=100$\kms. 
		\label{fig:valley_intensity}}
\end{figure}
